A Plasma Torus Around a Young Low-Mass Star

### 2025-03-11 paper/ms.tex
update

$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & XXX  &           & 3 \\

@@

$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.6 \pm X.X$   & X \\

@@

$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   &  XXX             & X \\

@@

6: Derived quantity.

@@

{\it Viewing Orientation}---We fitted the rotational broadening of the
K~I (7700\,\AA) resonance line using the broadening kernel suggested by
Ref.~\cite{Gray2008}; taking the mean and standard deviation of the
result value over all epochs yielded $v_{\rm eq} \sin
i$=138$\pm$8\,\kms, consistent with the visual line broadening $\Delta
\lambda$$\approx$3\,\AA.  The star's equatorial velocity $v_{\rm eq}$
based on its apparent size and rotation period is 130$\pm$4\,\kms.
While this suggests that the viewing orientation could be nearly
edge-on, the formal constraint is rather weak, with $i$$>$63$^\circ$ at
2$\sigma$.
% from calc_max_amplitude_sinusoid_rv_timeseries.py
% from K_to_msini.py, assuming Mstar = 0.18

@@

{\it Age: No Obvious Association Membership}---Previous work
\cite{Bouma2024} has found that over 90\% of CPVs are associated with
known young moving groups based on their positions and kinematics.
TIC~141146667 is one of the exceptions.  We calculated the probability
of TIC~141146667 being part of any of the nearby known groups using
BANYAN\,$\Sigma$ v1.2 \cite{Gagne2018}.  Based on that model,
it appears to be a field star at $>$99.9\% confidence.  We also searched the
local vicinity of TIC~141146667 for neighbors with similar projected on-sky
velocities using \texttt{comove} \cite{Tofflemire2021}.  This yielded
no strong candidates for co-moving stars --- defined as those with
projected tangential velocities $\Delta v_{\rm T}$$<$5\,\kms ---
that share its isochronal youth.

@@

{\it Effective temperature, radius, and mass}
We adopt the results from the spectral energy distribution fitting
exercise described by \cite{Bouma2024}.
In brief, this approach used
\texttt{astroARIADNE}...
\cite{Vines2022} with the BT-Settl stellar
atmosphere models \cite{Allard2012} assuming the
\cite{Asplund2009} solar abundances, and the
\cite{Barber2006} water line lists.  
We used broadband magnitudes from Gaia DR2, APASS, 2MASS, SDSS, and WISE
  $W1$ and $W2$.  (explain)...

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry} \\ 
    \hline
    $\alpha$ & todo & Gaia\ DR3 \\
    $\delta$ & todo & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag) & todo & TESS\ \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Kinematics and Position} \\
    \hline
    $RV_{Bary}$ (km s$^{-1}$ ) & $13.35 \pm 3.39$ & HIRES \\
    $U$ (\kms) & & \\
    $V$ (\kms) & & \\
    $W$ (\kms) & & \\
    $X$ (pc) & -28.4 & \\
    $Y$ (pc) &  19.8 & \\
    $Z$ (pc) &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $P_{rot}$ (hours) & $3.930 \pm 0.XXX$ & This work \\ 
    $v \sin i_\star$(km s$^{-1}$) & $138 \pm 8$ & This work\\
    $i_\star$($^\circ$) & $>$63 & This work \\
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024}\\
    %$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    $L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & todo & This work\\
    $t_{\rm iso}$ (Myr) & $16^{+19}_{-6}$ &  This work \\
    \hline
    \end{tabular}
    \caption{Properties of \starname.}
    \label{tab:stellarParameters}
\end{table}

>>>>

$M_\star$\dotfill                       & Stellar mass ($M_\odot$)\dotfill                               & 0.19$\pm$0.01  & 3 \\

@@

$d$\dotfill                             & Distance (pc)\dotfill                                          & $57.54 \pm 0.09$   & 7 \\

@@

$R_{\rm c}$\dotfill		                  & Keplerian corotation\dotfill                                   & $1.72 \pm 0.06$  & 6 \\

@@

6: Derived quantity,
7: Gaia DR3 geometric \cite{GaiaDR3}.

@@

{\it Viewing Orientation}---We fitted the rotational broadening of the
K~I (7700\,\AA) resonance line using the broadening kernel suggested by
\cite{Gray2008}; taking the mean and standard deviation of the
resulting value over all epochs yielded $v_{\rm eq} \sin
i$=138$\pm$8\,\kms, consistent with the visual line broadening $\Delta
\lambda$$\approx$3\,\AA.  The star's equatorial velocity $v_{\rm eq}$
based on its apparent size and rotation period is 130$\pm$4\,\kms.
While this suggests that the viewing orientation could be nearly
edge-on, the formal constraint is rather weak, with $i$$>$63$^\circ$ at
2$\sigma$ (2.5$^{\rm th}$ percentile of the inclination posterior).
% from calc_max_amplitude_sinusoid_rv_timeseries.py
% from K_to_msini.py, assuming Mstar = 0.18

@@

{\it Age: No Obvious Association Membership}---Previous work
\cite{Bouma2024} has found that over 90\% of CPVs are associated with
known young moving groups based on their positions and kinematics.
TIC~141146667 is one of the exceptions.  We calculated the probability
of TIC~141146667 being part of any of the nearby known groups using
BANYAN\,$\Sigma$ v1.2 \cite{Gagne2018}.  That particular model
classifies it as a field star at $>$99.9\% confidence.  We also searched the
local vicinity of TIC~141146667 for neighbors with similar projected on-sky
velocities using \texttt{comove} \cite{Tofflemire2021}.  This yielded
no strong candidates for co-moving stars --- defined as those with
projected tangential velocities $\Delta v_{\rm T}$$<$5\,\kms ---
that share its isochronal youth.

@@

{\it Effective temperature, radius, and mass}
For the stellar effective temperature and radius, we adopted the results
from the spectral energy distribution fitting methods described by
\cite{Bouma2024}.
In brief, this approach used \texttt{astroARIADNE} \cite{Vines2022} with
the BT-Settl stellar atmosphere models \cite{Allard2012}, assuming the
\cite{Asplund2009} solar abundances, and the \cite{Barber2006} water
line lists.  
This routine fitted for the stellar effective temperature, radius,
reddening, surface gravity, metallicity, and distance by comparing the
measured broadband magnitudes against pre-computed model grids.
Specifically, we used the broadband magnitudes from Gaia DR2, APASS,
2MASS, SDSS, and WISE $W1$ and $W2$.
This method has the best constraining power for the star's effective
temperature and radius; the surface gravity, metallicity, and reddening
are typically only weakly constrained.
Given the effective temperature, stellar radius, and previously
determined age, we then estimated the stellar mass by interpolating
against the PARSEC v1.2S isochrones \cite{Chen2014}. 
Specifically, we used the distance metric defined in Equation~8 of
\cite{Bouma2024}, in order to select the model mass closest to the
observed temperature, radius, and age.  We derived the uncertainty in
the mass based by propagating the statistical uncertainties in each of
these parameters.

@@

% TEMPLATE: IRAS041
%
\begin{table}
    \centering
    \begin{tabular}{lcr}
    \hline 
    \hline
    Parameter & Host & Source \\
    \hline 
    \multicolumn{3}{c}{Identifiers} \\
    \hline
    TIC & 141146667 & TESS \\
    Gaia & 860453786736413568 & Gaia\ DR3 \\
    %2MASS & J04154278+2909597 & J04154269+2909558 & 2MASS \\
    %ALLWISE & J041542.77+290959.5 & ... & ALLWISE\\
    \hline
    \multicolumn{3}{c}{Astrometry} \\ 
    \hline
    $\alpha$ & todo & Gaia\ DR3 \\
    $\delta$ & todo & Gaia\ DR3 \\
    $\mu_{\alpha}$ (mas yr$^{-1}$ ) & -73.933 $\pm$ 0.022 & Gaia\ DR3 \\
    $\mu_{\delta}$ (mas yr$^{-1}$ ) &  32.262 $\pm$ 0.024 & Gaia\ DR3 \\
    $\pi$ (mas)                     &  17.324 $\pm$ 0.025 & Gaia\ DR3 \\
    RUWE                            &  1.23               & Gaia\ DR3 \\
    \hline
    \multicolumn{3}{c}{Photometry} \\
    \hline
    $TESS$ (mag) & todo & TESS\ \\
    $G$ (mag)                       & 14.701 $\pm$ 0.002 & Gaia\ DR3 \\
    $G_{\rm BP}$ (mag)              & 16.664 $\pm$ 0.008 & Gaia\ DR3 \\
    $G_{\rm RP}$ (mag)              & 13.398 $\pm$ 0.006 & Gaia\ DR3 \\
    $G_{\rm BP}$-$G_{\rm RP}$ (mag) &  3.276 $\pm$ 0.010 & Gaia\ DR3 \\
    $J$ (mag)                       & 11.401 $\pm$ 0.022 & 2MASS     \\
    $H$ (mag)                       & 10.793 $\pm$ 0.021 & 2MASS     \\
    $K_s$ (mag)                     & 10.473 $\pm$ 0.016 & 2MASS     \\
    $W1$ (mag)                      & 10.276 $\pm$ 0.023 & ALLWISE   \\ % Cutri+2013 II/328/allwise
    $W2$ (mag)                      & 10.070 $\pm$ 0.020 & ALLWISE   \\
    $W3$ (mag)                      &  9.838 $\pm$ 0.045 & ALLWISE   \\
    %$W4$ (mag)                      &  8.392 $\pm$ NOUNC & ALLWISE \\
	  %
    %TODO: eROSITA?  ROSAT is an emitter.  GALEX detection.
    % Ofek+2011 1.4Ghz emitter.  Helfand+2015 ditto.  Bruzewski+2021.
    \hline
    \multicolumn{3}{c}{Kinematics and Position} \\
    \hline
    $RV_{Bary}$ (km s$^{-1}$ ) & $13.35 \pm 3.39$ & HIRES \\
    $U$ (\kms) & & \\
    $V$ (\kms) & & \\
    $W$ (\kms) & & \\
    $X$ (pc) & -28.4 & \\
    $Y$ (pc) &  19.8 & \\
    $Z$ (pc) &  67.0 & \\
    \hline
    \multicolumn{3}{c}{Physical Properties} \\
    \hline
    $P_{rot}$ (hours) & $3.930 \pm 0.XXX$ & This work \\ 
    $v \sin i_\star$(km s$^{-1}$) & $138 \pm 8$ & This work\\
    $i_\star$($^\circ$) & $>$63 & This work \\
    $T_{eff}$ (K) & 2972 $\pm$ 40 & \cite{Bouma2024}\\
    %$\log{g_{\star}}$\dotfill              & Surface Gravity (cgs)\hspace{9pt}\dotfill                      & YYY              & X \\
    $R_\star$ ($R_{\odot}$) & 0.42 $\pm$ 0.02 & \cite{Bouma2024} \\
    $A_V$ (mag) & 0 & \cite{Green2019} \\
    $F_{bol}$ (erg cm$^{-2}$ s$^{-1}$ ) & todo & This work\\
    $L_\star$ ($L_{\odot}$)  & todo & This work\\
    $M_\star$ ($M_{\odot}$)  & 0.188 $\pm$ 0.012  & This work\\
    $t_{\rm iso}$ (Myr) & $16^{+19}_{-6}$ &  This work \\
    \hline
    \end{tabular}
    \caption{Properties of \starname.}
    \label{tab:stellarParameters}
\end{table}
